Offshore Blue Crab Densities for <br/> The Gulf of Mexico

### 2020-02-04 2019_edits/R/offshore_analyses/crabs_per_hectare.Rmd
effort and catch maps added

```{r}
year_sums <- seamap %>% 
  group_by(survey_year) %>% 
  summarise(Stations          = n(),
            `Missing Obs`     = sum(is.na(cpue_towspd)),
            `Mean Catch`      = mean(cpue_towspd, na.rm = T),
            `SD`              = sd(cpue_towspd, na.rm = T),
            `Var`             = var(cpue_towspd),
            `Overdispersed`   = ifelse(Var > `Mean Catch`, "yes", "no"),
            `Catch SE`        = sd(cpue_towspd, na.rm = T) / sqrt(Stations - `Missing Obs`),
            `% Occurrence`    = mean(crab_presence, na.rm = T),
            `% Occurrence SE` = sd(crab_presence, na.rm = T) / sqrt(Stations - `Missing Obs`)) %>% 
  ungroup() %>% 
    mutate(Survey_Year        = factor(survey_year),
           catch_ymin         = `Mean Catch` - (2 * `Catch SE`),
           catch_ymax         = `Mean Catch` + (2 * `Catch SE`),
           catch_ymin         = ifelse(catch_ymin < 0, 0, catch_ymin),
           occurrence_ymin    = `% Occurrence` - (2 * `% Occurrence SE`),
           occurrence_ymax    = `% Occurrence` + (2 * `% Occurrence SE`),
           occurrence_ymin    = ifelse(occurrence_ymin < 0, 0, occurrence_ymin)
         )

@@

year_foa <- year_sums %>% 
  mutate(survey_year = as.numeric(as.character(survey_year))) %>% 
  ggplot() +
    geom_col(aes(survey_year, `% Occurrence`)) +
    geom_errorbar(aes(x = survey_year, 
                      ymin = occurrence_ymin,
                      ymax = occurrence_ymax),
                  size = .25) +
  scale_y_continuous(labels = scales::percent, limits = c(0,1)) + 
  labs(x = NULL, y = "Percent Occurrence")

@@

```{r}
rs_summary <- seamap %>% 
  filter(is.na(region)      == FALSE) %>% 
  group_by(region, season) %>% 
  summarise(Stations          = n(),
            `Missing Obs`     = sum(is.na(cpue_towspd)),
            `Mean Catch`      = mean(cpue_towspd, na.rm = T),
            `SD`              = sd(cpue_towspd, na.rm = T),
            `Var`             = var(cpue_towspd),
            `Overdispersed`   = ifelse(Var > `Mean Catch`, "yes", "no"),
            `Catch SE`        = sd(cpue_towspd, na.rm = T) / sqrt(Stations - `Missing Obs`),
            `% Occurrence`    = mean(crab_presence, na.rm = T),
            `% Occurrence SE` = sd(crab_presence, na.rm = T) / sqrt(Stations - `Missing Obs`)) %>% 
    ungroup() %>% 
    mutate(Season             = factor(season, levels = c("Summer", "Fall")),
           region             = factor(region, 
                                       levels = c("Texas", "Louisiana", "MS Bight", "Florida")),
           #region             = fct_rev(region),
           catch_ymin         = `Mean Catch` - (2 * `Catch SE`),
           catch_ymax         = `Mean Catch` + (2 * `Catch SE`),
           catch_ymin         = ifelse(catch_ymin < 0, 0, catch_ymin),
           occurrence_ymin    = `% Occurrence` - (2 * `% Occurrence SE`),
           occurrence_ymax    = `% Occurrence` + (2 * `% Occurrence SE`),
           occurrence_ymin    = ifelse(occurrence_ymin < 0, 0, occurrence_ymin)
    )

#Summary Table
rs_summary %>% 
  select(Region = region, Season, Stations, `Missing Obs`, `% Occurrence`, `Mean Catch`, SD, Overdispersed) %>% 
  knitr::kable()

# Plot
rs_plot <- rs_summary %>% 
  ggplot(aes(region, `Mean Catch`, fill = season)) +
    geom_col(position = "dodge") +
    geom_errorbar(aes(ymin = catch_ymin,
                      ymax = catch_ymax),
                  position = "dodge", 
                  size = .25) +
  scale_fill_manual(name = "", values = setNames(c("gray35", "gray50"), c("Summer", "Fall"))) +
  labs(x = NULL,
       y = "Mean Catch (C. sapidus / Hectare)") + 
  theme(
    legend.position = c(1, .95),
    legend.justification = c("right", "top"),
    legend.box.just = "right",
    legend.margin = margin(0, 6, 2, 2),
    legend.background = element_rect(fill = NA))

@@

rs_plot_2 <- rs_summary %>% 
  ggplot(aes(region, `% Occurrence`, fill = Season)) +
    geom_col(position = "dodge") +
    geom_errorbar(aes(ymin = occurrence_ymin,
                      ymax = occurrence_ymax),
                  position = "dodge",
                  size = .25) +
  scale_fill_manual(name = "", values = setNames(c("gray35", "gray50"), c("Summer", "Fall"))) +
  scale_y_continuous(labels = scales::percent, limits = c(0,1)) + 
  labs(x = NULL,
       y = "Percent Occurrence") + theme(
    legend.position = c(1, .95),
    legend.justification = c("right", "top"),
    legend.box.just = "right",
    legend.margin = margin(0, 6, 2, 2),
    legend.background = element_rect(fill = NA))

@@

```{r}
season_summary <- seamap %>% 
  group_by(survey_year, season) %>% 
  summarise(Stations          = n(),
            `Missing Obs`     = sum(is.na(cpue_towspd)),
            `Mean Catch`      = mean(cpue_towspd, na.rm = T),
            `SD`              = sd(cpue_towspd, na.rm = T),
            `Var`             = var(cpue_towspd),
            `Overdispersed`   = ifelse(Var > `Mean Catch`, "yes", "no"),
            `Catch SE`        = sd(cpue_towspd, na.rm = T) / sqrt(Stations - `Missing Obs`),
            `% Occurrence`    = mean(crab_presence, na.rm = T),
            `% Occurrence SE` = sd(crab_presence, na.rm = T) / sqrt(Stations - `Missing Obs`)) %>% 
    ungroup() %>% 
    mutate(survey_year        = factor(survey_year),
           # StatZone         = factor(StatZone),
           # StatZone         = fct_rev(StatZone),
           catch_ymin         = `Mean Catch` - (2 * `Catch SE`),
           catch_ymax         = `Mean Catch` + (2 * `Catch SE`),
           catch_ymin         = ifelse(catch_ymin < 0, 0, catch_ymin),
           occurrence_ymin    = `% Occurrence` - (2 * `% Occurrence SE`),
           occurrence_ymax    = `% Occurrence` + (2 * `% Occurrence SE`),
           occurrence_ymin    = ifelse(occurrence_ymin < 0, 0, occurrence_ymin)
         )

season_summary %>% select(`Survey Year` = survey_year, Season = season, Stations, `Missing Obs`, `% Occurrence`, `Mean Catch`, SD, Overdispersed) %>% knitr::kable()

season_plot <- season_summary %>% 
  mutate(survey_year = as.numeric(as.character(survey_year))) %>% 
  ggplot(aes(survey_year, `Mean Catch`, fill = season)) +
    geom_col(position = "dodge") +
    geom_errorbar(aes(ymin = catch_ymin,
                      ymax = catch_ymax),
                  position = "dodge",
                  size = .25) +
  scale_fill_manual(name = "", values = setNames(c("gray35", "gray50"), c("Summer", "Fall"))) +
  labs(x = NULL,
       y = "Mean Catch (C. sapidus / Hectare)") + theme(
    legend.position = c(1, .95),
    legend.justification = c("right", "top"),
    legend.box.just = "right",
    legend.margin = margin(0, 6, 2, 2),
    legend.background = element_rect(fill = NA))

@@

season_plot_2 <- season_summary %>% 
  mutate(survey_year = as.numeric(as.character(survey_year))) %>% 
  ggplot(aes(survey_year, `% Occurrence`, fill = season)) +
    geom_col(position = "dodge") +
    geom_errorbar(aes(ymin = occurrence_ymin,
                      ymax = occurrence_ymax),
                  position = "dodge",
                  size = .25) +
  scale_fill_manual(name = "", values = setNames(c("gray35", "gray50"), c("Summer", "Fall"))) +
  scale_y_continuous(labels = scales::percent, limits = c(0,1)) + 
  labs(x = "Survey Year",
       y = "Percent Occurrence") + theme(
    legend.position = c(1, .95),
    legend.justification = c("right", "top"),
    legend.box.just = "right",
    legend.margin = margin(0, 6, 2, 2),
    legend.background = element_rect(fill = NA))

@@

>>>>

```{r}
year_sums <- seamap %>% 
  group_by(survey_year) %>% 
  summarise(Stations          = n(),
            `Missing Obs`     = sum(is.na(cpue_towspd)),
            `Mean Catch`      = mean(cpue_towspd, na.rm = T),
            `SD`              = sd(cpue_towspd, na.rm = T),
            `Var`             = var(cpue_towspd),
            `Overdispersed`   = ifelse(Var > `Mean Catch`, "yes", "no"),
            `Catch SE`        = sd(cpue_towspd, na.rm = T) / sqrt(Stations - `Missing Obs`),
            `% Occurrence`    = mean(crab_presence, na.rm = T),
            `% Occurrence SE` = sd(crab_presence, na.rm = T) / sqrt(Stations - `Missing Obs`)
            ) %>% 
  ungroup() %>% 
    mutate(Survey_Year        = factor(survey_year),
           catch_ymin         = `Mean Catch` - (2 * `Catch SE`),
           catch_ymax         = `Mean Catch` + (2 * `Catch SE`),
           catch_ymin         = ifelse(catch_ymin < 0, 0, catch_ymin),
           occurrence_ymin    = `% Occurrence` - (2 * `% Occurrence SE`),
           occurrence_ymax    = `% Occurrence` + (2 * `% Occurrence SE`),
           occurrence_ymin    = ifelse(occurrence_ymin < 0, 0, occurrence_ymin)
         )

@@

year_foa <- year_sums %>% 
  mutate(survey_year = as.numeric(as.character(survey_year))) %>% 
  ggplot() +
    geom_col(aes(survey_year, `% Occurrence`)) +
    # geom_errorbar(aes(x = survey_year, 
    #                   ymin = occurrence_ymin,
    #                   ymax = occurrence_ymax),
    #               size = .25) +
  scale_y_continuous(labels = scales::percent, limits = c(0,1)) + 
  labs(x = NULL, y = "Percent Occurrence")

@@

*NOTE 2: Mean Catch (per hectare) and frequency of occurrence calculated for each year, in each region, for each season. The mean and standard error of those annual measurements are then taken*

```{r old}
# rs_summary <- seamap %>% 
#   filter(is.na(region)      == FALSE) %>% 
#   group_by(region, season) %>% 
#   summarise(Stations          = n(),
#             `Missing Obs`     = sum(is.na(cpue_towspd)),
#             `Mean Catch`      = mean(cpue_towspd, na.rm = T),
#             `SD`              = sd(cpue_towspd, na.rm = T),
#             `Var`             = var(cpue_towspd),
#             `Overdispersed`   = ifelse(Var > `Mean Catch`, "yes", "no"),
#             `Catch SE`        = sd(cpue_towspd, na.rm = T) / sqrt(Stations - `Missing Obs`),
#             `% Occurrence`    = mean(crab_presence, na.rm = T),
#             `% Occurrence SE` = sd(crab_presence, na.rm = T) / sqrt(Stations - `Missing Obs`)) %>% 
#     ungroup() %>% 
#     mutate(Season             = factor(season, levels = c("Summer", "Fall")),
#            region             = factor(region, 
#                                        levels = c("Texas", "Louisiana", "MS Bight", "Florida")),
#            #region             = fct_rev(region),
#            catch_ymin         = `Mean Catch` - (2 * `Catch SE`),
#            catch_ymax         = `Mean Catch` + (2 * `Catch SE`),
#            catch_ymin         = ifelse(catch_ymin < 0, 0, catch_ymin),
#            occurrence_ymin    = `% Occurrence` - (2 * `% Occurrence SE`),
#            occurrence_ymax    = `% Occurrence` + (2 * `% Occurrence SE`),
#            occurrence_ymin    = ifelse(occurrence_ymin < 0, 0, occurrence_ymin)
#     )

# #Summary Table
# rs_summary %>% 
#   select(Region = region, Season, Stations, `Missing Obs`, `% Occurrence`, `Mean Catch`, SD, Overdispersed) %>% 
#   knitr::kable()

#Calaculate Annual average cpue and percent occurrence
rs_summary <- seamap %>% 
  group_by(survey_year, region, season) %>% 
  summarise(cpue = mean(cpue_towspd, na.rm = T),
            percent_occurrence = mean(crab_presence, na.rm = T)) %>% 
  group_by(region, season) %>% 
  summarise(nyears = n(),
            avg_annual_cpue = mean(cpue, na.rm = T),
            annual_cpue_se = sd(cpue, na.rm = T) / sqrt(nyears),
            avg_percent_occurrence = mean(percent_occurrence, na.rm = T),
            po_se = sd(percent_occurrence, na.rm = T) / sqrt(nyears)) %>% 
    ungroup() %>%
    mutate(Season             = factor(season, levels = c("Summer", "Fall")),
           region             = factor(region, levels = c("Texas", "Louisiana", 
                                                          "MS Bight", "Florida")),
           catch_ymin         = avg_annual_cpue - (2 * annual_cpue_se),
           catch_ymax         = avg_annual_cpue + (2 * annual_cpue_se),
           catch_ymin         = ifelse(catch_ymin < 0, 0, catch_ymin),
           occurrence_ymin    = avg_percent_occurrence - (2 * po_se),
           occurrence_ymax    = avg_percent_occurrence + (2 * po_se),
           occurrence_ymin    = ifelse(occurrence_ymin < 0, 0, occurrence_ymin),
           occurrence_ymax    = ifelse(occurrence_ymax > 1, 1, occurrence_ymax)
    )

# Plot
rs_plot <- rs_summary %>% 
  ggplot(aes(region, avg_annual_cpue, fill = season)) +
    geom_col(position = "dodge") +
    geom_errorbar(aes(ymin = catch_ymin,
                      ymax = catch_ymax),
                  position = "dodge", 
                  size = .25) +
  scale_fill_manual(name = "", values = setNames(c("gray35", "gray50"), c("Summer", "Fall"))) +
  labs(x = NULL,
       y = "Mean Catch (C. sapidus / Hectare)") + 
  theme(
    legend.position = c(1, .95),
    legend.justification = c("right", "top"),
    legend.box.just = "right",
    legend.margin = margin(0, 6, 2, 2),
    legend.background = element_rect(fill = NA))

@@

rs_plot_2 <- rs_summary %>% 
  ggplot(aes(region, avg_percent_occurrence, fill = Season)) +
    geom_col(position = "dodge") +
    geom_errorbar(aes(ymin = occurrence_ymin,
                      ymax = occurrence_ymax),
                  position = "dodge",
                  size = .25) +
  scale_fill_manual(name = "", values = setNames(c("gray35", "gray50"), c("Summer", "Fall"))) +
  scale_y_continuous(labels = scales::percent, limits = c(0,1)) + 
  labs(x = NULL,
       y = "Percent Occurrence") + theme(
    legend.position = c(1, .95),
    legend.justification = c("right", "top"),
    legend.box.just = "right",
    legend.margin = margin(0, 6, 2, 2),
    legend.background = element_rect(fill = NA))

@@

*NOTE 2: Mean Catch (per hectare) and frequency of occurrence calculated for each year, for each season. The mean and standard error of those annual measurements are then taken*

```{r}
# season_summary <- seamap %>% 
#   group_by(survey_year, season) %>% 
#   summarise(Stations          = n(),
#             `Missing Obs`     = sum(is.na(cpue_towspd)),
#             `Mean Catch`      = mean(cpue_towspd, na.rm = T),
#             `SD`              = sd(cpue_towspd, na.rm = T),
#             `Var`             = var(cpue_towspd),
#             `Overdispersed`   = ifelse(Var > `Mean Catch`, "yes", "no"),
#             `Catch SE`        = sd(cpue_towspd, na.rm = T) / sqrt(Stations - `Missing Obs`),
#             `% Occurrence`    = mean(crab_presence, na.rm = T),
#             `% Occurrence SE` = sd(crab_presence, na.rm = T) / sqrt(Stations - `Missing Obs`)) %>% 
#     ungroup() %>% 
#     mutate(survey_year        = factor(survey_year),
#            # StatZone         = factor(StatZone),
#            # StatZone         = fct_rev(StatZone),
#            catch_ymin         = `Mean Catch` - (2 * `Catch SE`),
#            catch_ymax         = `Mean Catch` + (2 * `Catch SE`),
#            catch_ymin         = ifelse(catch_ymin < 0, 0, catch_ymin),
#            occurrence_ymin    = `% Occurrence` - (2 * `% Occurrence SE`),
#            occurrence_ymax    = `% Occurrence` + (2 * `% Occurrence SE`),
#            occurrence_ymin    = ifelse(occurrence_ymin < 0, 0, occurrence_ymin)
#          )
# 
# season_summary %>% select(`Survey Year` = survey_year, Season = season, Stations, `Missing Obs`, `% Occurrence`, `Mean Catch`, SD, Overdispersed) %>% knitr::kable()

#Calaculate Annual average cpue and percent occurrence
season_summary <- seamap %>% 
  group_by(survey_year, season) %>% 
  summarise(cpue = mean(cpue_towspd, na.rm = T),
            percent_occurrence = mean(crab_presence, na.rm = T)) %>% 
  group_by(season) %>% 
  summarise(nyears = n(),
            avg_annual_cpue = mean(cpue, na.rm = T),
            annual_cpue_se = sd(cpue, na.rm = T) / sqrt(nyears),
            avg_percent_occurrence = mean(percent_occurrence, na.rm = T),
            po_se = sd(percent_occurrence, na.rm = T) / sqrt(nyears)) %>% 
    ungroup() %>%
    mutate(Season             = factor(season, levels = c("Summer", "Fall")),
           catch_ymin         = avg_annual_cpue - (2 * annual_cpue_se),
           catch_ymax         = avg_annual_cpue + (2 * annual_cpue_se),
           catch_ymin         = ifelse(catch_ymin < 0, 0, catch_ymin),
           occurrence_ymin    = avg_percent_occurrence - (2 * po_se),
           occurrence_ymax    = avg_percent_occurrence + (2 * po_se),
           occurrence_ymin    = ifelse(occurrence_ymin < 0, 0, occurrence_ymin),
           occurrence_ymax    = ifelse(occurrence_ymax > 1, 1, occurrence_ymax)
    )

season_plot <- season_summary %>% 
  ggplot(aes(season, avg_annual_cpue, fill = season)) +
    geom_col() +
    geom_errorbar(aes(ymin = catch_ymin,
                      ymax = catch_ymax),
                  size = .25) +
  scale_fill_manual(name = "", values = setNames(c("gray35", "gray50"), c("Summer", "Fall"))) +
  labs(x = NULL,
       y = "Mean Catch (C. sapidus / Hectare)") + theme(
    legend.position = c(1, .95),
    legend.justification = c("right", "top"),
    legend.box.just = "right",
    legend.margin = margin(0, 6, 2, 2),
    legend.background = element_rect(fill = NA))

@@

season_plot_2 <- season_summary %>% 
  ggplot(aes(season, avg_percent_occurrence, fill = season)) +
    geom_col(position = "dodge") +
    geom_errorbar(aes(ymin = occurrence_ymin,
                      ymax = occurrence_ymax),
                  size = .25) +
  scale_fill_manual(name = "", values = setNames(c("gray35", "gray50"), c("Summer", "Fall"))) +
  scale_y_continuous(labels = scales::percent, limits = c(0,1)) + 
  labs(x = NULL,
       y = "Percent Occurrence") + theme(
    legend.position = c(1, .95),
    legend.justification = c("right", "top"),
    legend.box.just = "right",
    legend.margin = margin(0, 6, 2, 2),
    legend.background = element_rect(fill = NA))

@@

## Spatial Grids  {.tabset}

library(stars)
library(sf)
library(rnaturalearth)

#State and Country Bounds
usmap <- ne_states(country = "united states of america") %>% 
  st_as_sf() %>% 
  filter(region == "South")

mex <- ne_states(country = "mexico") %>% 
  st_as_sf()

#Custom grid for rasterize
test_bbox <- st_bbox(c(xmin = -97.5, xmax = -81,
                       ymin = 24.3, ymax = 30.5),
                     crs = st_crs(4326))

num_x <- seq(from = -97.4, to = -81.3, by = 0.2)
num_y <- seq(from = 24.4, to = 30.4, by = 0.2)

####  Aggregation  ####
#Aggregating the seamap model into bins
seamap_sf_agg <- seamap %>% 
  mutate(lon_bin = round(start_long, 1),
         lat_bin = round(start_lat, 1)) %>% 
  group_by(lon_bin, lat_bin) %>% 
  summarise(n_stations = n(),
            total_catch = sum(sapidus_catch, na.rm = T),
            avg_cpue = mean(cpue_towspd)) %>% 
  st_as_sf(coords = c("lon_bin", "lat_bin"), crs = 4326, remove = FALSE)

####  Total Catch  ####

#Convert to Raster and plot 
catch_ras <- st_rasterize(
  seamap_sf_agg["total_catch"],
  template = st_as_stars(
      test_bbox,
      #st_bbox(seamap_sf), 
      values = NA_real_,
      nx = length(num_x),
      ny = length(num_y))
  )

#plot(catch_ras, axes = T, main = "Total Sapidus Catch - All Years 2000+")

####  Mean CPUE  ####

cpue_ras <- st_rasterize(
  seamap_sf_agg["avg_cpue"],
  template = st_as_stars(
      test_bbox,
      #st_bbox(seamap_sf), 
      values = NA_real_,
      nx = length(num_x),
      ny = length(num_y))
  )

#plot(cpue_ras, axes = T, main = " Mean Catch/Hectare - All Years 2000+")

####  Sampling Effort  ####
effort_ras <- st_rasterize(
  seamap_sf_agg["n_stations"],
  template = st_as_stars(
      test_bbox,
      #st_bbox(seamap_sf), 
      values = NA_real_,
      nx = length(num_x),
      ny = length(num_y))
  )

### Sampling Effort

```{r}
####  Sampling Effort  ####
ggplot() +
  geom_stars(data = effort_ras) +
  geom_sf(data = usmap) +
  geom_sf(data = mex) +
  scale_fill_distiller(palette = "Spectral", na.value = "transparent") +
  coord_sf(xlim = c(-98, -81), ylim = c(24, 31)) +
  labs(x = NULL, y = NULL) +
  guides(fill = guide_colorbar(title = "Total Sampling Effort (Stations)",
                               title.hjust = 0.5, 
                               title.position = "top",
                               barwidth = 10)) +
  theme(legend.position = "bottom")
```

### Catch per Hectare

####  CPUE  ####
ggplot() +
  geom_stars(data = cpue_ras) +
  geom_sf(data = usmap) +
  geom_sf(data = mex) +
  scale_fill_distiller(palette = "Spectral", na.value = "transparent") +
  coord_sf(xlim = c(-98, -81), ylim = c(24, 31)) +
  labs(x = NULL, y = NULL) +
  guides(fill = guide_colorbar(title = "Average Catch/Hectare",
                               title.hjust = 0.5, 
                               title.position = "top",
                               barwidth = 10)) +
  theme(legend.position = "bottom")

### Total Catch

```{r}
####  Total Catch  ####
ggplot() +
  geom_stars(data = catch_ras) +
  geom_sf(data = usmap) +
  geom_sf(data = mex) +
  scale_fill_distiller(palette = "Spectral", na.value = "transparent") +
  coord_sf(xlim = c(-98, -81), ylim = c(24, 31)) +
  labs(x = NULL, y = NULL) +
  guides(fill = guide_colorbar(title = "Total C. sapidus Caught",
                               title.hjust = 0.5, 
                               title.position = "top",
                               barwidth = 10)) +
  theme(legend.position = "bottom")
